---
title: Simpson Guest Stars - Interactive Tree
author: Vaibhav Singh
date: '2019-11-17'
slug: interactive-tree
categories:  ["Data Blogs"]
tags: ["Data Visualisation","Tree"]
toc: FALSE  
subtitle: "Analysis of a Simpson dataset - Source Wikipedia"
---


```{r global options, include = FALSE}
knitr::opts_chunk$set(echo=TRUE, include = TRUE, warning=FALSE, message=FALSE)
```
# Simpsons Guest Stars

## Before getting started
Seeking to get answers for below based on dataset:<br/>
1. Which Season did maximum guest stars feature in? -- refer plot \@ref(fig:1)<br/> 
2. Who were the guest stars with maximum appearance on the Simpsons? -- refer plot \@ref(fig:2)<br/>

## Lets get started

Read more about data set [here](https://github.com/rfordatascience/tidytuesday/tree/master/data/2019/2019-08-27)

```{r echo=FALSE, warning=FALSE}
# Load libraries
library(tidyverse)
library(collapsibleTree) 
library(htmlwidgets)
library(knitr)
library(printr)
theme_set(theme_minimal())
simpsons <- readr::read_delim("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-08-27/simpsons-guests.csv", delim = "|", quote = "") %>% mutate(self = str_detect(role, "self|selves"),
         season = parse_number(season))

```

Previewing the data
```{r echo=FALSE}
kable(head(simpsons))
```

1. Which Season did maximum guest stars feature in ?
```{r 1, fig.cap='It seems Season 20 onwards frequency of guest stars started increasing',message=FALSE, warning=FALSE, echo=FALSE}
simpsons %>% count(season) %>% 
  ggplot(aes(season,n))+
  geom_col(aes(fill=as.factor(season)))+
  coord_flip()+
  theme(legend.position="none")+
  labs(title="Which Season had the most guest stars",
       subtitle="Seems Seasons 20 onwards Simpsons started calling up more guest stars",
       x="Season Number",
       y="Count of Guest stars")

```

2. Which guest star had the most appearance on Simpsons
```{r 2,fig.cap='Marcia Wallace & Phil Hartman have the maximum guest appearance',message=FALSE, warning=FALSE, echo=FALSE}

grouped_data  <- simpsons %>% 
  separate_rows(role, sep = ";\\s+") %>% 
  na.omit() %>% 
  mutate(number=parse_number(number)) %>% 
  count(season,guest_star,sort=TRUE) %>% 
  group_by(season) %>%
  top_n(1,wt=n) %>% 
  mutate(percent=round(n/sum(n),1))  %>% 
  inner_join(simpsons %>% count(guest_star,sort=TRUE) %>% head(30),by=c("guest_star"))
  
grouped_data$Color <- grouped_data$guest_star
levels(grouped_data$Color) <- colorspace::rainbow_hcl(11)

h<-collapsibleTree(
    grouped_data,
    root = "SIMPSONS", 
    hierarchy = c( "guest_star","season" ,"n.x"), 
    nodeSize = "percent", 
    width = 1000, 
    zoomable = TRUE
 )

h
```